Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-12 Final Draft/Final Draft.Rmd
final draft updates

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(rticles)
library(tidyverse)
library(stargazer)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(readxl)
library(scales)
library(tigris)
library(dplyr)
library(CBPS)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")

@@

```
\FloatBarrier
This coefficient is still somewhat difficult to interpret, and made increasingly so by the winner-take-all system by which U.S. Electoral College votes are allocated. As an alternative perspective, we examined a range of counterfactual scenarios, in which counties manufacturing experienced job losses between the 1st and 99th percentiles across all counties. For example, in the map below, we assume that all counties in the US experienced a loss in white manufacturing employment equivalent to the level experienced by a county at the 25th percentile. Here, that level is equivalent to a 0 net loss in white manufacturing jobs between 2004-2015 per 100 workers in 2004 (a relatively low level of job losses). Setting all counties' white manufacturing job losses to 0, we use the point estimates of our regression coefficients from Model 3 in \ref{regResult04} to estimate the counterfactual change in Democratic vote share. We then aggregate our county-level findings to estimate the overall counterfactual change in Democratic vote share at the state level. We find that, had all counties experienced the 25th percentile level in manufacturing job losses, most states would have leaned more Democratic in 2016, and that Michigan and Pennsylvania would have flipped.

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

@@

\caption{Counterfactual Presidential Election Simulation Results}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

# first plot CI
ggplot(summary, aes(ptile, median_result)) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                color = "blue") +
  geom_hline(yintercept = 270, color = "red") +
  theme_bw() +
  ylab("Clinton Electoral College Votes\n") +
  xlab("Assumed Percentile of Manufacturing Layoffs\n")   +
  theme(axis.title.x = element_text(size = 8), 
        axis.title.y = element_text(size = 8))

\FloatBarrier
\begin{figure} \label{counterfactual3}
\caption{ Counterfactual Clinton Win Frequency}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

ggplot(summary2, aes(ptile, perc_wins)) + 
  geom_col(fill = "#4287f5")  +
  theme_bw() +
  ylab("Clinton Wins\nIn 1000 Simulations") +
  xlab("Percentile of Manufacturing Layoffs\n") +
  ggtitle("Counterfactual Presidential Election Simulation Results")

>>>>

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(rticles)
library(tidyverse)
library(stargazer)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(readxl)
library(scales)
library(tigris)
library(dplyr)
library(glmnet)
library(CBPS)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")

@@

```
\FloatBarrier
This coefficient is still somewhat difficult to interpret, and made increasingly so by the winner-take-all system by which U.S. Electoral College votes are allocated. As an alternative perspective, we examined a range of counterfactual scenarios, in which counties manufacturing experienced job losses between the 1st and 99th percentiles across all counties.

For example, in the map in Figure \ref{counterfactual1} below, we assume that all counties in the US experienced a loss in white manufacturing employment equivalent to the level experienced by a county at the 25th percentile. Here, that level is equivalent to a 0 net loss in white manufacturing jobs between 2004-2015 per 100 workers in 2004 (a relatively low level of job losses). Setting all counties' white manufacturing job losses to 0, we use the point estimates of our regression coefficients from Model 4 in Table \ref{regResult04} to estimate the counterfactual change in Democratic vote share. We find that, had all counties experienced the 25th percentile level in manufacturing job losses, most states would have leaned more Democratic in 2016, and that Michigan and Pennsylvania would have flipped.

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Counterfactual Scenario: Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

We also demonstrate uncertainity around our coefficient estimates using a simulation appraoch. Specifically, for each percentile of white manufacturing job losses, we simulate 1,000 election outcomes by drawing from the distribution around our coefficent point estimate 1,000 times (again, using the results from Model 3 in Table \ref{regResult04}). In Figure \ref{counterfactual2}, we find that, assuming the 24th percentile or lower of white manufacturing job losses, we find that the median estimate in our simulation shows Democratic winning at least 270 electoral college votes. Using the 95\% confidence intervals produced from the simulation exercise, we find that we would be 95\% confident that Democrats would have won the national election had white manufacturing job losses been at the 5th percentile or lower. This result suggest that manufacturing job losses would have had to have been substantially lower and *negative* (e.g., job gains) in order for the national election outcome to have been flipped with 95\% confidence.

@@

\caption{Counterfactual Simulation Results: Democratic Electoral College Votes Assuming White Manufacturing Job Losses Between 1st and 99th Percentile }
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

# first plot CI
ggplot(summary, aes(ptile, median_result)) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                color = "blue") +
  geom_hline(yintercept = 270, color = "red") +
  theme_bw() +
  ylab("Democratic Electoral College Votes\n") +
  xlab("Assumed Percentile of White Manufacturing Job Losses\n")   +
  theme(axis.title.x = element_text(size = 8), 
        axis.title.y = element_text(size = 8))

```
\begin{figurenotes}
Figure shows a 95\% confidence interval for Democratic Electoral College Votes at Each Percentile of Manufacturing Job Losses. The black dot indicates the median electoral college vote estimate at each percentile. The red line shows the 270 electoral vote threshold.
\end{figurenotes}
\end{figure}
\FloatBarrier

Finally, using the same simulation approach described above, in Figure \ref{counterfactual3} we illustrate the freuqency with which Demcorats would have won the presidential election (270 votes) given each percentile of white manufacturing job losses. Our simulation results show that, when we assumed 24th percentile or lower of white manufacturing job losses,  Democrats won the presidential election in at least 50 percent of simulations. When we assumed the 17th percentile or lower, Democrats won the presidential election in at least 75 percent of simulations. Taken together, these results suggest that Democrats would have had a significantly higher chance of winning the presidential election had white manufacturing job layoffs been substantially lower.

\FloatBarrier
\begin{figure} \label{counterfactual3}
\caption{Counterfactual Simulation Results: Democratic Presidential Win Frequency}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

ggplot(summary2, aes(ptile, perc_wins)) + 
  geom_col(fill = "#4287f5")  +
  theme_bw() +
  ylab("Frequency of Democratic Presidential Win\nIn 1000 Simulations") +
  xlab("Assumed Percentile of White Manufacturing Job Losses\n") +
  ggtitle("") + theme(text = element_text(size = 8))

### 2021-12-12 Final Draft/Final-Draft.tex
final draft updates

Like \cite{Baccini21}, we find evidence that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. \cite{Baccini21} find that gross
manufacturing layoffs between 2012-2015 were associated with a loss in
Democratic vote share. However, using a different measure of
deindustrialization - \emph{net} change in manufacturing employment - we
find the oppposite effect: that counties with manufacturing job
\emph{gains} from 2012-2015 tended to swing towards Donald Trump in
2016. We extend the analysis to the 2004-2015 period in order to capture
the long-term effects of deindustrialization on a region. Doing so, we
find that regions experiencing more manufacturing job losses swung
towards Donald Trump in 2016, and that this occurred more when those
losing jobs were white.

@@

@@

\FloatBarrier
\end{figure}
\begin{figurenotes}
Figures show a 95% confidence interval for Democratic Electoral College Votes at Each Percentile of Manufacturing Job Losses.
\end{figurenotes}
\FloatBarrier

\FloatBarrier
\begin{figure} \label{counterfactual3}
\caption{Clinton Winning Freq}

>>>>

@@

We also demonstrate uncertainity around our coefficient estimates using
a simulation appraoch. Specifically, for each percentile of white
manufacturing job losses, we simulate 1,000 election outcomes by drawing
from the distribution around our coefficent point estimate (again, using
the results from Model 3 in Table \ref{regResult04}). In Figure
\ref{counterfactual2}, we find that, assuming the 24th percentile or
lower of white manufacturing job losses, we find that the median
estimate in our simulation shows Democratic winning at least 270
electoral college votes. Using the 95\% confidence intervals produced
from the simulation exercise, we find that we would only be 95\%
confident that Democrats would have won the national election had white
manufacturing job losses been at the 5th percentile or lower. This
result suggest that manufacturing job losses would have had to have been
\emph{negative} (e.g., job gains) in order for the national election
outcome to have been flipped with 95\% confidence.

@@

\FloatBarrier
\begin{figurenotes}
Figure shows a 95\% confidence interval for Democratic Electoral College Votes at Each Percentile of Manufacturing Job Losses. The black dot indicates the median electoral college vote estimate at each percentile. The red line shows the 270 electoral vote threshold.
\end{figurenotes}
\end{figure}
\FloatBarrier

Finally, using the same simulation approach described above, in Figure
\ref{counterfactual3} we illustrate the freuqency with which Demcorats
would have won the presidential election (270 votes) given each
percentile of white manufacturing job losses. Our simulation results
show that, when we assumed 24th percentile or lower of white
manufacturing job losses, Democrats won the presidential election in at
least 50 percent of simulations. When we assumed the 17th percentile or
lower, Democrats won the presidential election in at least 75 percent of
simulations. Taken together, these results suggest that Democrats would
have had a significantly higher chance of winning the presidential
election had white manufacturing job layoffs been substantially lower.

\FloatBarrier
\begin{figure} \label{counterfactual3}
\caption{Counterfactual Simulation Results: Democratic Presidential Win Frequency}
